---
title       : Comparative Investing - Delta Hedge Fund vs Savings Account
subtitle    : Market vs Best Available Bank Rate
author      : B.A. McDougall
job         : NSCI Consulting
framework   : io2012     # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Portfolio Management

### Generate Risk Neutral Valued (RNV) Portfolio using [Binary Tree Model](https://en.wikipedia.org/wiki/Binomial_options_pricing_model)
* Purchasing $\Delta$ shares of stock at price $S_o$, and short a call option on same stock with value $f$
* Value of RNV portfolio is same whether stock price goes up or down; therefore, $(\Delta S_o) u - f_u = (\Delta S_o) d - f_d$.  $u$ is multiplier for increase; $d$ is multiplier for decrease.

### $\Delta$ Hedge for RNV Portfolio
* From equality, $\Delta$ required for creating a RNV Portfolio is $\Delta = \frac{f_u - f_d}{S_o (u-d)}$
* PV of RNV is discounted FV $\Longrightarrow (\Delta S_o)  - f = e^{-r t} \left[ (\Delta S_o) u - f_u \right]$
* Solving for f estimates PV of call option $\Longrightarrow f = e^{-r t} \left[ p f_u + (1 - p) f_d \right]$
where
<div class="centered">
\begin{equation}
p = \frac{e^{r t} - d}{u - d}
\end{equation}
</div>

* **NOTE**:  f is discounted expectation value of option's FV against a binomial density (*coin flip*)

---
## Portfolio Management (continued)
* For stochastic processes, variance is proportional to $\Delta t$, so
<div class="centered">
\begin{equation}
\sigma^2 \Delta t = \langle S(t)^2 \rangle - \langle S(t) \rangle ^2
\end{equation}
</div>
<div class="centered">
\begin{equation}
\sigma^2 \Delta t = \left[ p u^2 + (1-p) d^2 \right] - \left[ \left( p u + (1-p) d \right)^2 \right]
\end{equation}
</div>
<div class="centered">
\begin{equation}
\sigma^2 \Delta t = (u + d) e^{\mu t} - du - e^{2 \mu t}
\end{equation}
</div>

* Using a Taylor Expansion, one validates a solution for u & d is $u = e^{\sigma \sqrt{\Delta t}}$ & $d = e^{- \sigma \sqrt{\Delta t}}$
* Summarizing:  (1) FV of RNV Portfolio is $\left[(\Delta S_o)  - f\right]e^{\mu t}$ which is of same form as FV of Savings account; (2) The relationship between $f$ and $S_o$ is fully determined.

## Pricing a Stock
* Two assumptions: (1) during a small time interval, a random contribution to the differential of stock price is $\epsilon \sqrt(\Delta t)$ where $\epsilon$ is a standard normal distribution; (2) Values of $\epsilon \sqrt(\Delta t)$ from any two time intervals $\Delta t$ are mutually independent.

---
## Time Dependence of $S$

* The differential of stock price is $d S = \mu S d t + \sigma S \epsilon \sqrt{dt}$. Neglecting volatility, $S(t) = S_o e^{\mu t}$.

* Solving numerically, $\Delta S = \mu S \Delta t + \sigma S \epsilon \sqrt{\Delta t}$, so $S_{i + 1} = S_i + \mu S_i \Delta t + \sigma S_i \epsilon \sqrt{\Delta t}$

```{r libraryLoad, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
# require(stringr)
# require(DiagrammeR)
require(slidify)
require(knitr)
require(ggplot2)
# require(grid)
# require(gridExtra)
# require(gtable)
require(lubridate)
library(dplyr)
# library(Cairo)
library(animation)
library(htmlwidgets)
ani.options(interval=.6)
```

```{r fvMoney, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=12, fig.cap="Time Value of Money"}
aprAnn = 0.0125 #https://www.mysavingsdirect.com/MySavingsDirectWeb/en/common/information/LearnMore.jsp
aprDay = aprAnn / yday(mdy("12/31/2015"))
aprMnth = aprAnn / 12
aprQtr = aprAnn / 4

finGrowthTblMnth <- tbl_df(as.data.frame( ymd_hms("2015-01-01 16:00:00", tz = "America/New_York") + months(1:120) - days(1) ))
names(finGrowthTblMnth) <- "dateClose"
finGrowthTblMnth <- finGrowthTblMnth%>%
    dplyr::mutate(
        ydayNum = yday(dateClose),
        month = month(dateClose),
        years = as.period( (dateClose - ymd_hms("2015-01-02 16:00:00", tz = "America/New_York") ) ) %/% months(12),
        quarters = as.period( (dateClose - ymd_hms("2015-01-02 16:00:00", tz = "America/New_York") ) ) %/% months(3),
        months = as.period( (dateClose - ymd_hms("2015-01-02 16:00:00", tz = "America/New_York") ) ) %/% months(1),
        days = as.period( (dateClose - ymd_hms("2015-01-02 16:00:00", tz = "America/New_York") ) ) %/% days(1),
        Savings = (532.60 * (1 + aprDay)**(days)),
        Contin = (532.60 * exp( aprDay * days) ),
        Mnthly = (532.60 * (1 + aprMnth)**(months)),
        Qtrly = (532.60 * (1 + aprQtr)**(quarters)),
        Annul = (532.60 * (1 + aprAnn)**(years))
        )

# mnyPlt2 <- ggplot() + scale_colour_hue("Compound_Rate")
# mnyPlt2 <- mnyPlt2 + geom_line(data=finGrowthTblMnth, aes(x = dateClose, y = Contin, color = "1_Contin"))
# mnyPlt2 <- mnyPlt2 + geom_line(data=finGrowthTblMnth, aes(x = dateClose, y = Savings, color = "2_Daily"))
# mnyPlt2 <- mnyPlt2 + geom_line(data=finGrowthTblMnth, aes(x = dateClose, y = Mnthly, color = "3_Mnthly"))
# mnyPlt2 <- mnyPlt2 + geom_line(data=finGrowthTblMnth, aes(x = dateClose, y = Qtrly, color = "4_Qtrly"))
# mnyPlt2 <- mnyPlt2 + geom_line(data=finGrowthTblMnth, aes(x = dateClose, y = Annul, color = "5_Annul"))
# mnyPlt2 <- mnyPlt2 + labs(x = "Date", y = "Amount [$]", title = "Money:  Time Value", legend = "Compounding")
# mnyPlt2
```

```{r stockPricingPrep, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
################################################################################
##
## This section is for comparing Google Stock Price with Wiener Pricing of Stock
##  For comparison, bank savings is shown.  Import Google price history. 
##
mu <- (547.47 - 532.60) / 532.60 # ~0.15 annualized rate of return from
# https://www.stock-analysis-on.net/NASDAQ/Company/Google-Inc/Valuation/Ratios
mu_daily <- mu / 112

finGrowthTbl <- tbl_df(
    read.csv2("googl.csv", sep = ",", stringsAsFactors = FALSE, na.strings = "NA",
              header = TRUE, #colClasses = c("character", "numeric", "numeric", "numeric", "numeric", "numeric")
              )
    )   %>%
    select(
        Date, Open, High, Low, Close, Volume
        ) %>%
    mutate(
        dateClose = ymd_hms( paste( dmy(Date), hms(c("16:00:00") )), tz = "America/New_York"),
        Open = as.numeric(Open), High = as.numeric(High), Low = as.numeric(Low), Close = as.numeric(Close),
        Volume = as.numeric(Volume)
        ) %>%
    select(
        dateClose,
        Date, Open, High, Low, Close, Volume      
        )  %>%
    filter(
        dateClose > mdy("01/01/2015") & dateClose < mdy("06/30/2015")
        ) %>%
    mutate(
        ydayNum = yday(dateClose),
        days = as.period( (dateClose - ymd_hms("2015-01-02 16:00:00", tz = "America/New_York") ) ) %/% days(1)        
        )%>%
    arrange(ydayNum)%>%
    mutate(
        Savings = (532.60 * (1 + aprDay)**(days)),
        Contin = (532.60 * exp( aprDay * days) ),
        goog_W = 0,
        goog_H = (532.60 * exp( mu_daily * days) )
        )%>%
    select(
        dateClose, days,
        Open, High, Low, Close, goog_W, goog_H, Savings
        )
##
################################################################################
##
## This section for determining volatility in Google Stock between 2015-01-02
##  and 2015-06-12, which is 112 trading days and estimate Stock price with Wiener model
##
sd_googl <- sd(finGrowthTbl$Close) # for 112 trading days
sd_googlDly <-  sd_googl / 112

goog_W <- vector("numeric", length = 112)
goog_W[1] <- 532.60
```

```{r stockPricing, echo=TRUE, message=TRUE, warning=TRUE, error=TRUE, cache=FALSE}
## This code chunk demonstrates pricing Google Stock assuming a Wiener process.
for(j in 1:30) {
    for(i in 2:112) {
        #     set.seed(314159)
        goog_W[i] <- goog_W[i-1] * (1 + mu_daily * (1/365) + sd_googlDly * rnorm(1) * sqrt(1/365)) 
        }
    }
print(paste('On 12 June 2015, the simulated closing price of Google Stock is $', round(goog_W[112],2)))
```
* Code chunk numerically integrates the opening Google stock price on 2 Jan 2015 of $532.60 through 112 trading days until 12 June 2015

---
## Animation of Pricing a Stock Using a Wiener Process
* Numerical simulation and actual stock price of Google (Dates:  1/2/2015 - 6/12/2015).
* Plot also shows savings account (best interest $r=1.25\%$) & a $\Delta$ Hedge with Google stock (market rate $\mu= 6.28 \%$ &
$ \sigma = $ \$41.)

```{r stockPricingAnimating, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.height=5, fig.width=12, fig.show='animate'}
################################################################################
##
## This section is for comparing Google Stock Price with Wiener Pricing of Stock
##  For comparison, bank savings is shown.  Import Google price history. 
##
for(j in 1:30) {
    for(i in 2:112) {
        #     set.seed(314159)
        goog_W[i] <- goog_W[i-1] * (1 + mu_daily * (1/365) + sd_googlDly * rnorm(1) * sqrt(1/365)) 
        }
    finGrowthTbl$goog_W <- goog_W
    
    mnyPlt <- ggplot() + scale_colour_hue("Instrument") + scale_y_continuous(limits = c(450, 650))
    mnyPlt <- mnyPlt + geom_line(data=finGrowthTbl, aes(x = dateClose, y = Close, color="GOOGL"))
    mnyPlt <- mnyPlt + geom_line(data=finGrowthTbl, aes(x = dateClose, y = goog_H, color = "goog_H"))
    mnyPlt <- mnyPlt + geom_line(data=finGrowthTbl, aes(x = dateClose, y = Savings, color = "Savings"))
    mnyPlt <- mnyPlt + labs(x = "Date", y = "Amount [$]", title = "Money:  Time Value")
    mnyPlt <- mnyPlt + geom_line(data=finGrowthTbl, aes(x = dateClose, y = goog_W, color="goog_W"))
    print(mnyPlt)
    }
```
* Animation demonstrates stochastic nature of stock pricing
